File-based localization of numerical perturbations in data analysis pipelines

### 2020-07-29 paper.tex
Answers to some review comments

%%%% Commands %%%%
% \newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
% \newcommand{\note}[2]{\color{blue}Note from \textsc{#1}: #2\color{black}}
\newcommand{\revision}[1]{\color{purple}\textbf{REVISION:}#1\color{black}}
\newcommand{\revised}[1]{\color{blue}\textbf{REVISED:}#1\color{black}}
\newcommand{\reprozip}[0]{ReproZip\xspace}
\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
\newcommand{\toolname}[0]{Spot\xspace}
\newcommand{\flirt}[0]{\texttt{flirt}\xspace}
\newcommand{\fnirt}[0]{\texttt{fnirt}\xspace}

@@

Vibrations in computational infrastructures impact data
analyses in various fields, but identifying the origin of these effects in complex pipelines
remains challenging.  In some cases, most likely due to numerical
instabilities, small perturbations resulting from changes in operating
system versions~\cite{Glatard2015}, hardware~\cite{jezequel2015estimation}, or parallelization parameters~\cite{diethelm2011limits},
may result in substantially different analysis outcomes. To better understand and 
correct these effects, efficient tools are 
needed to assist pipeline developers
in the comparison of results obtained across different conditions.

@@

>>>>

%%%% Commands %%%%
% \newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
% \newcommand{\note}[2]{\color{blue}Note from \textsc{#1}: #2\color{black}}
\newcommand{\revision}[1]{\color{purple}\textbf{REVISION:}#1\color{black}}
\newcommand{\revised}[1]{\color{blue}#1\color{black}\xspace}
\newcommand{\reprozip}[0]{ReproZip\xspace}
\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
\newcommand{\toolname}[0]{Spot\xspace}
\newcommand{\flirt}[0]{\texttt{flirt}\xspace}
\newcommand{\fnirt}[0]{\texttt{fnirt}\xspace}

@@

\revised{Numerical perturbations resulting from variations in computational
environments} impact data analyses in various fields, but identifying the
origin of these perturbations in complex pipelines remains challenging.  In
some cases, small perturbations resulting from changes in operating system
versions~\cite{Glatard2015}, hardware~\cite{jezequel2015estimation}, or
parallelization parameters~\cite{diethelm2011limits}, result in
substantially different analysis outcomes, \revised{due to the propagation
and amplification of floating-point errors. While the existence of such
numerical errors is well known~\cite{stoer2013introduction}, their impact
on scientific computations has multiplied with the rise of the Big Data
era, due to the sustained growth of data sets, the increasing complexity of
analysis pipelines, and the diversification of computing infrastructures.}
To better understand and 
correct these effects, efficient tools are 
needed to assist pipeline developers
in the comparison of results obtained across different conditions.

@@

\revised{
Computational environments are only one of many factors contributing to the
on-going reproducibility crisis. In fact, sample size selection,
publication bias, or methodological flexibility in the analysis are likely
to have a stronger effect than numerical perturbations, although
to our knowledge no evidence of this is available. We refer to the
studies in~\cite{botvinik2020variability,bowring2019exploring,bhagwat2020understanding,kennedy2019everything} for deeper
analyses of the associated effects on neuroimaging analyses. It should also be
noted that the effects of computational environments and these other
factors manifest at different levels: referring to the terminology used
in~\cite{peng2011reproducible}, computational environments are associated with
\emph{reproducibility}, the minimal standard by which identical results
should be obtainable from identical data and parameters, while the other
aforementioned factors belong to \emph{replicability}, the ultimate
standard by which independent experimenters should be able to draw similar
conclusions from similar experiments. In practice, variability resulting
from computational environments manifests during software testing (test
results depend on execution platform), deployment on HPC systems (results
obtained on local vs HPC systems differ), or software version updates
(results obtained before vs after the update differ), while factors related
to replicability impact the community more broadly. Ultimately, both
reproducibility and replicability should be understood and improved.
}
